---
title: "jstadden_5"
author: "Jared Stadden"
date: "4/19/2020"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown


```{r}
#Read data into R
original = read.csv("C:\\Users\\jared\\Desktop\\Cereals.csv")

cereal <- na.omit(original)

#install.packages("stats")
#install.packages("cluster")
#install.packages("factoextra")

library(stats)
library(cluster)
library(ISLR)

```
.

```{r}
#Normalize data
cereal[,4:16] <- scale(cereal[,4:16])
df <- cereal

#Applying various methods
hc_single <- agnes(df, method = "single")
hc_complete <- agnes(df, method = "complete")
hc_average <- agnes(df, method = "average")
hc_ward <- agnes(df, method = "ward")

#Choose method with output closest to 1
print(hc_single$ac)
print(hc_complete$ac)
print(hc_average$ac)
print(hc_ward$ac)
```

1. Ward method is best since its value is closest to 1.


```{r}
#Applying Ward Heirarchical-Clustering
d <- dist(df, method = "euclidean")
hc_ward <- hclust(d, method = "ward")  

#Plot Dendrogram
plot(hc_ward, cex = 0.6)

#Dendrogram with rectangles for each cluster
rect.hclust(hc_ward, k = 4, border = 1:4)
```
2. Based on the dendrogram I am going to use 4 clusters


```{r}
#Assigning clusters
memb <- cutree(hc_ward, k=4)
df$cluster <- memb
```

```{r}
#Splitting the data
set.seed(123)
Index_Train<- sample(seq_len(nrow(cereal)),size=floor(0.6*nrow(cereal)))
Train<-cereal[Index_Train,]
Validation<-cereal[-Index_Train,]

#Clustering
d2 <- dist(Train, method = "euclidean")
hc_ward2 <- hclust(d2, method = "ward")  
memb <- cutree(hc_ward2, k=4)
Train$cluster <- memb
head(Train)
```
3. Couldn't get centroids to work



4. The data should be normalized. When data is not normalized the distance calculations of the clustering method place extra importance on variables with larger values. When assessing the healthiness of cereal we wouldn't want potassium to be overvalued compared to more important indicators such as vitamins, sugars, or fat. Normalization would ensure that clusters are created more fairly across measures of health with different units and scales but equal importance. 